Data-driven Reduction of Multiscale Stochastic Dynamical Systems \thanks{This work was supported by ....}

### 2014-09-02 reducible_SDE/paper/SIADS_draft.tex
added more to paper

Assuming that the functions $a_i$ are bounded and Lipschitz, and we can find
\begin{equation}
\mathbb{E} \left| \frac{1}{T} \int_t^{t + T} a_i(x_1, \dots, x_m, X_{m+1}(s), \dots, X_n(s)) ds - \overline{a} (x_1, \dots, x_m) \right| < \kappa_i (T) 
\end{equation}
for $1 \le i \le m$, where $\kappa_i((T) \rightarrow 0 $ as $T \rightarrow \infty$, then the averaging principle \cite{...} states that we can write an effective equation in the slow variables $x_1, \dots, x_m$. 
%
We will consider systems that can be written as deterministic (perhaps nonlinear) bi-Lipshitz functions $\mathbf{f}: \mathbb{R}^n \mapsto \mathbb{R}^d$ of the SDE in \eqref{eq:general_SDE}.
%
Our goal is to recover the slow variables $x_1, \dots, x_m$ using data-driven techniques.

@@

To locally recover the slow variables, we need to build a metric that is invariant to the fast variables. 
%
Typically, this is done by averaging out the fast variables. 
%
We propose to do this using the Mahalanobis distance, which is a local PCA-type approach that inverts the local covariance matrix of the data, thereby creating a space with normalized diffusion terms.
%
It can be shown that the Mahalanobis distance locally approximates: 
\begin{equation} \label{eq:mahalanobis}
\| \mathbf{f}(\Xi_1) - \mathbf{f}(\Xi_2) \|^2_M = \| \Xi_1 - \Xi_2 \|^2_2 + \mathcal{O}(\| \mathbf{f}(\Xi_1) - \mathbf{f}(\Xi_2) \|^4_2)
\end{equation}
where the components of $\Xi$ have uncoupled noise with unit variance,
i.e., the variables $\xi_1, \dots, \xi_n$ are governed by the following SDEs:
\begin{equation} \label{eq:NIV_formulation}
d\xi_i = b(\xi) dt + dW_i
\end{equation}
and $\Xi = \begin{bmatrix} \xi_1 & \xi_2& \cdots & \xi_n \end{bmatrix}^T$.

In particular, for SDEs of the form in \eqref{eq:general_SDE}, consider the transformation
\begin{equation} \label{eq:general_rescale}
\begin{aligned}
x_i = \xi_i, \: &1 \le i \le m \\
x_i = \frac{\xi_i}{\sqrt{\epsilon}} \: &1 \le i \le m
\end{aligned}
\end{equation}

@@

\begin{figure}
\epsfig{width=\textwidth, file=schematic.eps}
\caption{Illustration of how to choose $\delta t$ and $\sigma_{kernel}$ appropriately. The data shows the evolution of the ``fast'' variable at a fixed value of the ``slow'' variable.  We must choose both parameters so that the curvature effects and other nonlinearities are negligable. }
\label{fig:schematic}
\end{figure}

Our approximation of the pairwise distances will rely on the Taylor expansion of the measurement function $f$. 
%
We will then {\em empirically} estimate the first-order term in this expansion using simulation bursts to estimate the local covariance.  
%
Accordingly, we will address the accuracy of our method from two standpoints: (1) the accuracy of the Taylor expansion, and (2) the accuracy of the covariance estimation (see Figure~\ref{fig:schematic}).
%
We will present both analytical results for the error bounds, as well as an empirical methodology to set the appropriate parameters for our method to accurately recover the intrinsic slow variable(s).

@@

Given the true covariance, we can write the Euclidean distance $\|X_2 - X_1 \|$ using Taylor expansion as...

Talk about fourth-order error terms, say we want to choose the kernel scale such that these error terms are negligible. 
%
The error terms involve curvature of $f$, etc.

Let $\mathbf{g} = \mathbf{f}^{-1}: \mathbb{R}^d \mapsto \mathbb{R}^n$, and let $Y_1 = f(X_1)$, and $Y_2 = f(X_2)$.
%
Again, we let $g_j^i = \frac{\partial g^i}{\partial Y^j}$.
%
Then, by Taylor expansion, we obtain
\begin{equation}
\begin{aligned}
x_{2,i} &=& 
x_{1,i} 
+ \sum_{j=1}^d \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} (y_{2,j} - y_{1,j} ) 
+ \frac{1}{2} \sum_{j=1}^d \sum_{k=1}^d \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) \\
&&+ \frac{1}{6} \sum_{j=1}^d \sum_{k=1}^d \sum_{l=1}^d \left. \frac{\partial^3 g_i}{\partial y_j \partial y_k \partial y_l} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) (y_{2,l} - y_{1,l})
+ \mathcal{O}( \|Y_2 - Y_1\|^4 )
\end{aligned}
\end{equation}

Substituting the Taylor expansion into the quadratic form, we obtain
\begin{equation}
\begin{aligned}
\| X_2 - X_1 \|^2_2 
=& \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J J^T)^{-1}\right|_{Y_1} + \left.(J J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 ) \\
& + \frac{1}{2} \sum_{i=1}^n \sum_{jkl=1}^{d} \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_1} - \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l}) \\
& + \frac{1}{8} \sum_{i=1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \frac{1}{6}  \sum_{i=1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{aligned}
\end{equation}

In general, we do not have access to $f$, $g$, or any of its derivatives.
%
However, we note that $J J^T(Y_1)$ is the local covariance of $Y_1$, which we can empirically estimate from data (this will be discussed in a subsequent section).
%
Therefore, we choose to truncate the distance approximation at the first order term.
%
We call this distance the {\em Mahalanobis distance}, 
\begin{equation}
 \| Y_2 - Y_1 \|^2_M = \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J J^T)^{-1}\right|_{Y_1} + \left.(J J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 )
\end{equation}
%
The error in the approximation is defined as
\begin{equation}
e_M(Y_1, Y_2) = \| X_2 - X_1 \|^2_2 - \| Y_2 - Y_1 \|^2_M 
\end{equation}

We can bound the error by
\begin{equation}
| e_M(Y_1, Y_2)  | \le n \left( \left| K_1 K_2 \right| + \left| \frac{ K_2^2}{4} \right|  + \left| \frac{K_1 K_3}{3} \right|  \right) \| Y_2 - Y_1 \| ^4  
+ \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{equation}
%
where
%
\begin{equation}
\begin{aligned}
K_1 &= \sup_{i,j,Y} |g_j^i(Y)|\\
K_2 &= \sup_{i,j,k,Y} |g_{jk}^i(Y)|\\
K_3 &= \sup_{i,j,k,l,Y} |g_{jkl}^i(Y)|
\end{aligned}
\end{equation}

@@

To compute the Mahalanobis distance, we require $J J^T$. 
%
From \cite{...}, we know that $JJ^T = C$, where $C$ is the local covariance of the observed stochastic process. 
%
Therefore, we would like to estimate the local covariance at a point $X_t$.

@@

Let $X_\tau$ denote the samples at time $\tau > t$. 

@@

@@

@@

@@

e_C (X_t, \tau - t) = \| \hat{C} (X_t, \tau - t)  - C(X_t, \tau - t) \|

@@

@@

@@

In the first example, our transformation $\mathbf{f}$ will be the identity function and fast and slow remain uncoupled,
\begin{equation}
\mathbf{f}(x_1, x_2) = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
\end{equation}
%
and
%
\begin{equation}
\mathbf{g}(y_1, y_2) = \mathbf{f}^{-1} (y_1, y_2) = \begin{bmatrix} y_1 \\ y_2 \end{bmatrix}
\end{equation}

@@

\begin{equation}
\begin{aligned}
e_{C, jk} (X_t, \tau - t) = & 
\frac{1}{\tau - t} \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\tau - t} \sum_{i=1}^m  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \mathcal{O} ((\tau - t)^{3/2})
 \end{aligned}
\end{equation}

@@

In the second example, our data will be warped into half-moon shapes
\begin{equation}
\mathbf{f_2}(x_1, x_2) = 
\begin{bmatrix} 
x_1 + x_2^2 \\
x_2
\end{bmatrix}
\end{equation}

>>>>

Assuming that the functions $a_i$ are bounded and Lipschitz, and we can find
\begin{equation}
\mathbb{E} \left| \frac{1}{T} \int_t^{t + T} a_i(x_1, \dots, x_m, x_{m+1}(s), \dots, x_n(s)) ds - \overline{a} (x_1, \dots, x_m) \right| < \kappa_i (T) 
\end{equation}
for $1 \le i \le m$, where $\kappa_i((T) \rightarrow 0 $ as $T \rightarrow \infty$, then the averaging principle \cite{...} states that we can write an effective equation in the slow variables $x_1, \dots, x_m$.

We will consider systems that can be written as deterministic (perhaps nonlinear) bi-Lipshitz functions $\mathbf{f}: \mathbb{R}^n \mapsto \mathbb{R}^d$ of the SDE in \eqref{eq:general_SDE}.
%
Our goal is to recover the slow variables $x_1, \dots, x_m$  from observations of $\mathbf{f}(x_1, \dots, x_n)$ using data-driven techniques.

@@

To locally recover the slow variables, we need to build a metric that is invariant to the fast variables. 
%
Typically, this is done by averaging out the fast variables. 
%
We propose to do this using the Mahalanobis distance, which is a local PCA-type approach that inverts the local covariance matrix of the data, thereby creating a space with normalized diffusion terms.
%
It can be shown that the Mahalanobis distance locally approximates: 
\begin{equation} \label{eq:mahalanobis}
\| \mathbf{f}(X_1) - \mathbf{f}(X_2) \|^2_M = \| \Xi_1 - \Xi_2 \|^2_2 + \mathcal{O}(\| \mathbf{f}(X_1) - \mathbf{f}(X_2) \|^4_2)
\end{equation}
where the components of $\Xi$ are sampled from a stochastic process with uncoupled noise with unit variance,
i.e., $\Xi = \begin{bmatrix} \xi_1 & \xi_2& \cdots & \xi_n \end{bmatrix}^T$, where $\xi_1, \dots, \xi_n$ are governed by the following SDEs:
\begin{equation} \label{eq:NIV_formulation}
d\xi_i = b(\xi) dt + dW_i
\end{equation}

In particular, for SDEs of the form in \eqref{eq:general_SDE}, consider the transformation
\begin{equation} \label{eq:general_rescale}
\begin{aligned}
x_i = \xi_i, \: &1 \le i \le m \\
x_i = \frac{\xi_i}{\sqrt{\epsilon}}, \: &m+1 \le i \le n
\end{aligned}
\end{equation}

@@

\begin{figure}[t]
\epsfig{width=\textwidth, file=schematic.eps}
\caption{Illustration of how to choose $\delta t$ and $\sigma_{kernel}$ appropriately. The data shows the evolution of the ``fast'' variable at a fixed value of the ``slow'' variable.  We must choose both parameters so that the curvature effects and other nonlinearities are negligable. }
\label{fig:schematic}
\end{figure}

Our approximation of the pairwise distances will rely on the Taylor expansion of the measurement function $\mathbf{f}$. 
%
We will then {\em empirically} estimate the first-order term in this expansion using simulation bursts to estimate the local covariance.  
%
Accordingly, we will address the accuracy of our method from two standpoints: (1) the accuracy of the Taylor expansion, and (2) the accuracy of the covariance estimation (see Figure~\ref{fig:schematic}).
%
We will present both analytical results for the error bounds, as well as an empirical methodology to set the appropriate parameters for our method to accurately recover the intrinsic slow variable(s).

@@

We want to calculate the distance $\|\Xi_2 - \Xi_1\|$ in terms of the observations $\mathbf{f}(X)$. 
Let $\mathbf{g} = \mathbf{f}^{-1}: \mathbb{R}^d \mapsto \mathbb{R}^n$.
By \eqref{eq:general_rescale}, we have
g_i(Y) = x_i = \xi_i,& \: 1 \le i \le m \\
\sqrt{\epsilon} g_i(Y) = \sqrt{\epsilon} x_i = \xi_i,& \: m+1 \le i \le n
Let $Y_1 = \mathbf{f}(X_1)$, and $Y_2 = \mathbf{f}(X_2)$.
Then, by Taylor expansion, we obtain
\xi_{2,i} &=& 
\xi_{1,i} 
+ \sum_{j=1}^d \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} (y_{2,j} - y_{1,j} ) 
+ \frac{1}{2} \sum_{j=1}^d \sum_{k=1}^d \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) \\
&&+ \frac{1}{6} \sum_{j=1}^d \sum_{k=1}^d \sum_{l=1}^d \left. \frac{\partial^3 g_i}{\partial y_j \partial y_k \partial y_l} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) (y_{2,l} - y_{1,l})
+ \mathcal{O}( \|Y_2 - Y_1\|^4 )
for $1 \le i \le m$, and 
\xi_{2,i} &=& 
\xi_{1,i} 
+ \sqrt{\epsilon} \left[ \sum_{j=1}^d \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} (y_{2,j} - y_{1,j} ) 
+ \frac{1}{2} \sum_{j=1}^d \sum_{k=1}^d \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) \right. \\
&& \left.+ \frac{1}{6} \sum_{j=1}^d \sum_{k=1}^d \sum_{l=1}^d \left. \frac{\partial^3 g_i}{\partial y_j \partial y_k \partial y_l} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) (y_{2,l} - y_{1,l}) \right]
+ \mathcal{O}( \|Y_2 - Y_1\|^4 )
for $m+1 \le i \le n$. 
Substituting the Taylor expansion into the quadratic form, we obtain
\begin{equation} \label{eq:distance_taylor_expansion}
&\| \Xi_2 - \Xi_1 \|^2_2 = \\
& \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J I_\epsilon J^T)^{-1}\right|_{Y_1} + \left.( J I_\epsilon J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 ) \\
& + \frac{1}{2} \sum_{i=1}^m \sum_{jkl=1}^{d} \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_1} - \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l}) \\
& + \frac{\epsilon}{2} \sum_{i=m+1}^n \sum_{jkl=1}^{d} \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_1} - \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l}) \\
& + \frac{1}{8} \sum_{i=1}^m \sum_{jklm=1}^d  \left( \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \frac{\epsilon}{8} \sum_{i=m+1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \frac{1}{6}  \sum_{i=1}^m \sum_{jklm=1}^d  \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \frac{\epsilon}{6}  \sum_{i=m+1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
where $J$ is the Jacobian of $\mathbf{f}$, and $I_\epsilon$ is a diagonal matrix with $I_{\epsilon,ii} = 1$ for $1 \le i \le m$, and $I_{\epsilon,ii} = \frac{1}{\epsilon}$ for $m+1 \le i \le n$.

In general, we do not have access to $\mathbf{f}$, $\mathbf{g}$, or any of its derivatives.
%
However, we note that $\left. J I_\epsilon J^T\right|_Y$ is the local covariance of the stochastic process at $Y$, which we can empirically estimate from data (this will be discussed in a subsequent section).
%
Therefore, we choose to truncate the distance approximation at the first order term.
%
We call this distance the {\em Mahalanobis distance}, 
\begin{equation}
 \| Y_2 - Y_1 \|^2_M = \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J I_\epsilon J^T)^{-1}\right|_{Y_1} + \left.(J I_\epsilon J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 )
\end{equation}
%
The error in the approximation is defined as
\begin{equation} \label{eq:mahanaobis_error}
e_M(Y_1, Y_2) = \| \Xi_2 - \Xi_1 \|^2_2 - \| Y_2 - Y_1 \|^2_M 
\end{equation}

%We can bound the error by
%\begin{equation}
%| e_M(Y_1, Y_2)  | \le n \left( \left| K_1 K_2 \right| + \left| \frac{ K_2^2}{4} \right|  + \left| \frac{K_1 K_3}{3} \right|  \right) \| Y_2 - Y_1 \| ^4  
%+ \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
%\end{equation}
%%
%where
%%
%\begin{equation}
%\begin{aligned}
%K_1 &= \sup_{i,j,Y} |g_j^i(Y)|\\
%K_2 &= \sup_{i,j,k,Y} |g_{jk}^i(Y)|\\
%K_3 &= \sup_{i,j,k,l,Y} |g_{jkl}^i(Y)|
%\end{aligned}
%\end{equation}

@@

To compute the Mahalanobis distance, we require $J I_\epsilon J^T$. 
%
From \cite{...}, we know that $J I_\epsilon J^T = C$, where $C$ is the local covariance of the observed stochastic process. 
%
Therefore, we would like to estimate the local covariance at a point $X_t$.

@@

Let $X_\tau$ denote samples at time $\tau > t$. 

@@

\begin{equation} \label{eq:estimated_cov}

@@

\begin{equation} \label{eq:estimated_cov_linear_f}

@@

\begin{equation} \label{eq:cov_error}

@@

e_C (X_t, \tau - t) = \hat{C} (X_t, \tau - t)  - C(X_t, \tau - t) 

@@

\section{Empirical Estimation of Parameters}

Given data governed by an SDE as in \eqref{eq:general_SDE}, we would like to empirically set the simulation and analysis parameters.
%
There are two parameters: $\sigma_{kernel}$, which is the diffusion maps kernel scale, and $\tau - t$, which is the timescale of the ``burst'' used to estimate the covariance.

$\sigma_{kernel}$ controls which distances we ``trust''.
%
Therefore, we want the Mahalanobis distance approximation to be accurate within a ball of radius $\sigma_{kernel}$. 
%
From \eqref{eq:distance_taylor_expansion}, we can see that, for $e_M$ to be small, we want $\|Y_2 - Y_1 \|^2_M$ to vary linearly with $\|Y_2 - Y_1\|^2_2$.

For $e_C$ to be accurate, we want $\hat{C}$ to not vary as a function of $\tau - t$.

We can use simulations to check both of these conditions empirically.

@@

For this example, the analytical covariance is
 \begin{equation}
C(X_t) = 
\begin{bmatrix}
1 & 0 \\
0 & \frac{1}{\epsilon}
\end{bmatrix}
\end{equation}

@@

In the first example, our transformation $\mathbf{f}$ will be the identity function and fast and slow remain uncoupled,
\begin{equation}
\mathbf{f}(x_1, x_2) = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
\end{equation}
%
\begin{equation}
\mathbf{g}(y_1, y_2) = \mathbf{f}^{-1} (y_1, y_2) = \begin{bmatrix} y_1 \\ y_2 \end{bmatrix}
\end{equation}

@@

%\begin{equation}
%\begin{aligned}
%e_{C, jk} (X_t, \tau - t) = & 
%\frac{1}{\tau - t} \left. \frac{\partial f_j}{\partial x_1} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^1 \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \frac{\partial a_1}{\partial x_1} \frac{\partial f_k}{\partial x_1}  + \frac{1}{\epsilon} \frac{\partial a_2}{\partial x_1} \frac{\partial f_k}{\partial x_2} \right)  dW_{s_1}^1 ds_2 \right) \right] \\ 
% & + \frac{1}{\epsilon (\tau - t)} \left. \frac{\partial f_j}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^2 \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(\frac{\partial a_1}{\partial x_2} \frac{\partial f_k}{\partial x_1}  + \frac{1}{\epsilon} \frac{\partial a_2 }{\partial x_2} \frac{\partial f_k}{\partial x_2} \right)  dW_{s_1}^2 ds_2 \right) \right] \\ 
% & + \frac{1}{\tau - t}  \left. \frac{\partial f_k}{\partial x_1} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^1 \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \frac{\partial a_1}{\partial x_1} \frac{\partial f_j}{\partial x_1}  + \frac{1}{\epsilon} \frac{\partial a_2}{\partial x_1} \frac{\partial f_j}{\partial x_2} \right) dW_{s_1}^1 ds_2 \right) \right] \\
%& + \frac{1}{\epsilon (\tau - t)}  \left. \frac{\partial f_k}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^2 \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \frac{\partial a_1}{\partial x_2} \frac{\partial f_j}{\partial x_1}  + \frac{1}{\epsilon} \frac{\partial a_2}{\partial x_2} \frac{\partial f_j}{\partial x_2} \right) dW_{s_1}^2 ds_2 \right) \right] \\
%& + \mathcal{O} ((\tau - t)^{3/2})\\
% = & 
%\frac{1}{\epsilon^2 (\tau - t)} \left. \frac{\partial f_j}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^2 \right)  \left(\int_t^\tau \int_{t}^{s_2} -\frac{\partial f_k}{\partial x_2} dW_{s_1}^2 ds_2 \right) \right] \\ 
%& + \frac{1}{\epsilon^2 (\tau - t)}  \left. \frac{\partial f_k}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^2 \right)  \left(\int_t^\tau \int_{t}^{s_2} -\frac{\partial f_j}{\partial x_2} dW_{s_1}^2 ds_2 \right) \right] \\
%& + \mathcal{O} ((\tau - t)^{3/2})
% \end{aligned}
%\end{equation}

From \eqref{eq:estimated_cov_linear_f}, we find that 
\begin{equation}
\begin{aligned}
e_{C, 11}(X_t, \tau - t) &=& 
e_{C, 12}(X_t, \tau - t) = 
e_{C, 21}(X_t, \tau - t) = 0 \\
e_{C, 22}(X_t, \tau - t) &=& 
\frac{-2}{\epsilon^2 (\tau - t)} \mathbb{E} \left[ \left(\int_t^\tau dW_s^2 \right)  \left(\int_t^\tau \int_{t}^{s_2} -dW_{s_1}^2 ds_2 \right) \right]  
+ \mathcal{O} ((\tau - t)^{3/2})
\end{aligned}
\end{equation}
%
Therefore, $\|e_C \| = \mathcal{O}\left(\frac{\tau - t}{\epsilon^2} \right)$ and so we need to choose $\tau - t < \epsilon^2$.

@@

In the second example, our data will be warped into half-moon shapes
\begin{equation}
\mathbf{f}(x_1, x_2) = 
\begin{bmatrix} 
x_1 + x_2^2 \\
x_2
\end{bmatrix}
\end{equation}

\begin{equation}
\mathbf{g}(y_1, y_2) = \mathbf{f}^{-1} (y_1, y_2) = \begin{bmatrix} y_1 - y_2^2 \\ y_2 \end{bmatrix}
\end{equation}

From \eqref{eq:distance_taylor_expansion}, we have 
\begin{equation}
e_M(Y_1, Y_2) = (\epsilon-1) (y_{2,2} - y_{1,2})^4 
+ \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{equation}

%Therefore, we have
%\begin{equation}
%\begin{aligned}
%\frac{\partial f_1}{\partial x_1} &=& 1 \\
%\frac{\partial f_1}{\partial x_2} &=& 2 x_2 \\
%\frac{\partial f_2}{\partial x_1} &=& 0 \\
%\frac{\partial f_2}{\partial x_2} &=& 1 \\
%\frac{\partial^2 f_1}{\partial x_1^2} = \frac{\partial^2 f_1}{\partial x_1 \partial x_2} &=& 0 \\
%\frac{\partial^2 f_1}{\partial x_2^2} &=& 2 \\
%\frac{\partial^2 f_2}{\partial x_1^2} = \frac{\partial^2 f_2}{\partial x_1 \partial x_2} = \frac{\partial^2 f_2}{\partial x_2^2} &=& 0 \\
%\end{aligned}
%\end{equation}
%
%\begin{equation}
%\begin{aligned}
%e_{C, jk} (X_t, \tau - t) = & 
%\frac{1}{\tau - t} \left. \frac{\partial f_j}{\partial x_1} \right|_{X_t} \mathbb{E} \left[ \left( \int_t^\tau dW_s^1  \right) \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_k}{\partial x_1^2} dW_{s_1}^1 dW_{s_2}^1  \right) \right] \\
%& + \frac{1}{\epsilon^{3/2} (\tau - t)} \left. \frac{\partial f_j}{\partial x_2} \right|_{X_t} \mathbb{E} \left[ \left( \int_t^\tau dW_s^2  \right) \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_k}{\partial x_2^2} dW_{s_1}^2 dW_{s_2}^2  \right) \right] \\
%& + \frac{1}{\tau - t} \left. \frac{\partial f_k}{\partial x_1} \right|_{X_t}  \mathbb{E} \left[ \left( \int_t^\tau dW_s^1 \right)  \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_j}{\partial x_1^2}  dW_{s_1}^1 dW_{s_2}^1 \right) \right] \\
%& + \frac{1}{\epsilon^{3/2} (\tau - t)} \left. \frac{\partial f_k}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left( \int_t^\tau dW_s^2 \right)  \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_j}{\partial x_2^2}  dW_{s_1}^2 dW_{s_2}^2 \right) \right] \\
%&+ \mathcal{O} ((\tau - t) ) \\
%= & 
%\frac{1}{\epsilon^{3/2} (\tau - t)} \left. \frac{\partial f_j}{\partial x_2} \right|_{X_t} \mathbb{E} \left[ \left( \int_t^\tau dW_s^2  \right) \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_k}{\partial x_2^2} dW_{s_1}^2 dW_{s_2}^2  \right) \right] \\
%& + \frac{1}{\epsilon^{3/2} (\tau - t)} \left. \frac{\partial f_k}{\partial x_2} \right|_{X_t}  \mathbb{E} \left[ \left( \int_t^\tau dW_s^2 \right)  \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_j}{\partial x_2^2}  dW_{s_1}^2 dW_{s_2}^2 \right) \right] \\
%&+ \mathcal{O} ((\tau - t) ) 
%\end{aligned}
%\end{equation}
